Joint training of classification model and similarity model for low-resource text classification in chatbot

### 2020-09-09 chatbot_classification/aaai_2021/paper.tex
modified:   paper.tex   new file:   paper.ver06.pdf

  {todo: Add joint NER and classification problem here.}

@@

      F-scores  on  five  task-specific  datasets text classification in chatbot

@@

  \subsection{Implementation Details}
  For  our  SFC  systems,  we  use RoBERTa-large as the pretrained model for both
  classification  task  and  sentence-pair similarity task. We fine-tune
  our  SFC  system  on  our datasets with max sequence length of 128,
  learning  rate  of 1.5e-5 for RoBERTa-large pretrained model and learning rate
  of  5e-4  for  task-specific  layer  with polynomial decay and Adam optimizer.
  Besides,  we  set  the  max number of the epoch to 100 and save the best model on
  validation  set based on testing result in weighted F-score (todo: reference). 
  We report the F-score as the main evaluation measure for all experiments.

@@

  \subsection{Result Analysis}
  All results in F-score (todo: reference) on the five datasets with different models and configurations are
  showed in Table  \ref{tbe:table2}.

  The performance of each model is measured in F-score on the fixed
  test  set  for  each  dataset.  
  Overall,  two-stage  SFC systems in multi-task training deliver  moderate 
  improvements  over  corresponding baselines, and joint SFC deliver more
  significant and consistent improvements over the best baseline, RoBERTa based
  classification. In following content, we dive into more details.

  \subsubsection*{Is multi-task and joint training working?} 
  First, we compare the four  SFC  variants to analyze the improvement
  brought  by  multi-task  and  joint-model  structure.

  Comparing with training two-stage SFC with multi-task, training with only task
  1,  corresponding  to  sentence pair similarity model, degrades by 0.8 percent
  point  on average, and training with only task 2, corresponding to top-K based
  classification  task,  degrades  by  0.45  percent  point  on  average.  These
  degradation  indicates  multi-task training in two-stage SFC is helpful for the
  system  performance.  Besides, task 2 play a relatively more important role in
  multi-task learning, and this aligns with their optimal weight settings.

  Joint SFC consistently outperforms two-stage SFC on 5 diverse datasets by 0.95
  percents  on  average. This observation supports the idea that the joint model
  structure  can  alleviate  the  limitation  brought by the lack of interaction
  between  two  stages.  In  following analysis, we will focus on the comparison
  between joint SFC and the other baseline models.

  \subsubsection*{Is  the  fusion of a classification model and similarity model
  helpful?}  We  compare  the performance of joint SFC with classification based
  and similarity based baselines.

@@

  Regarding the  classification  models,  joint SFC
  outperforms    TextCNN    by   21.52   percent points,   ALBERT-xxlarge   by   7.99
  percent points, BERT-large  by  7.73 percent points, RoBERTa-base by 3.79 
  percent points, RoBERTa-large by 2.04 percents on average F-score over 5 datasets.  
  Especially, ALBERT.xxlarge based is much more  unstable  in our senario. Thus, 
  we did multiple run with different settings, e.g.,  learning rate and batch size 
  for each dataset, and  report the  best  result  here. Comparatively,  
  RoBERTa  based is  the  most stable and has the best performance among
  all baselines and all experiment settings.

@@

  \subsubsection*{Does joint SFC improve the classfication in stage 1?}
  {todo: content}

@@

  Our main
  focus  is  to  study  chatbot  building  in common situation during real-world
  application  where  only  a few sample sentences are available for each class.
  The  experimental  results  for  the  3  diverse  intention detection datasets
  (ClINC150,  BANKING77,  HWU64)  under  few-shot  settings, 5 to 20 samples per
  class,  indicate that our proposed joint SFC can achieve 2.47 percents average
  improvement   over  one  of  the  most  powerful  baseline  models,  which  is
  RoBERTa-large  classification  model.  Moreover,  the  improvement  in F-score
  becomes  more  and  more prominent with the decrease of data sample number per
  class.

@@

  We  conduct  experiments on 5 different settings of hyperparameters, which are
  the  number of candidate class number $K$ and the number of sentence pairs $P$
  for  each  class,  on  ITG  and Amazon-670K datasets to find the most suitable
  setting  of  these  2 hyperparameters. We control the total number of sentence
  pairs  sampled for each user query into about the same level of amount(50-60).
  Through  analyzing the experimenal results shown in Table \ref{tbe:table3}, we
  find  that the best setting of hyperparameters might be different for datasets
  of  different  data  size.  The  class label number of ITG and Amazon-670k are
  similar  in  our  experimental setting, which is around 250. However the total
  amount of data points of ITG is approximately 2 times larger than Amazon-670k.
  Therefore,  the  setting of $K=5$ and $P=10$ is good enough for ITG since most
  of  the  classes  contains  over 10 data points. However, for Amazon-670k, the
  amount  of  data  sample  for  each  class  is  scarce, thus, we may need more
  candidate  classes to sample enough effective sentence pairs for sentence-pair
  model  layer  of  SFC.  In this way, the setting of $K=10$ and $P=5$ becomes a
  good choice. However, in spite of the effect on evaluation performance brought
  by different settings of hyperparameters, joint SFC still steadily outperforms
  two-stage  SFC  by  0.41  percents  in  overall  average  F-score  on  ITG and
  Amazon-670k datasets.

>>>>

  Another   line   of  interesting  work  is  the  joint  training  of  sentence
  classification   and  NER  (todo:  reference).  If  both  the  city  and  date
  information  are  recognized,  then  they  are good indicators that this query
  might  be  booking  an  ticket. This work can also be combined in our proposed
  framework.

@@

      F1 scores  on  five  task-specific  datasets text classification in chatbot

@@

  \subsection{Implementation Details}
  For  our  SFC  systems,  we  use RoBERTa-large as the pretrained model for both
  classification  task  and  sentence-pair similarity task. We fine-tune
  our  SFC  system  on  our datasets with max sequence length of 128,
  learning  rate  of 1.5e-5 for RoBERTa-large pretrained model and learning rate
  of  5e-4  for  task-specific  layer  with polynomial decay and Adam optimizer.
  Besides,  we  set  the  max number of the epoch to 100 and save the best model on
  validation  set based on testing result in weighted F1 score (todo: reference). 
  We report the F1 score as the main evaluation measure for all experiments.

@@

  \subsection{Result Analysis}
  All results in F1 score (todo: reference) on the five datasets with different models and configurations are
  showed in Table  \ref{tbe:table2}.

  The performance of each model is measured in F1 score on the fixed
  test  set  for  each  dataset.  
  Overall,  two-stage  SFC systems in multi-task training deliver  moderate 
  improvements  over  corresponding baselines, and joint SFC deliver more
  significant and consistent improvements over the best baseline, RoBERTa based
  classification. In following content, we dive into more details.

  \subsubsection*{Are multi-task and joint training working?} 
  First, we compare the four  SFC  variants to analyze the improvement
  brought  by  multi-task  and  joint-model  structure.

  Comparing with training two-stage SFC with multi-task, training with only task
  1,  namely the sentence pair similarity model, degrades by 0.8 percent
  point  on average, and training with only task 2, namely the top-K based
  classification  task,  degrades  by  0.45  percent  point  on  average.  These
  degradation  indicates  multi-task training in two-stage SFC is helpful for the
  system  performance.  Besides, task 2 plays a relatively more important role in
  multi-task learning, and this aligns with their optimal weight settings.

  Joint SFC consistently outperforms two-stage SFC with multi-task training  
  on 5 diverse datasets by 0.95 percents  on  average. 
  This observation supports the idea that the joint model
  structure  can  alleviate  the  limitation  brought by the lack of interaction
  between  two  stages.  In  following analysis, we will focus on the comparison
  between joint SFC and the other baseline models.

  \subsubsection*{Is  the  fusion of a classification model and a similarity model
  helpful?}  We  compare  the performance of joint SFC with classification based
  and similarity based baselines.

@@

  Regarding the  classification  models,  joint SFC
  outperforms    TextCNN    by   21.52   percent points,   ALBERT-xxlarge   by   7.99
  percent points, BERT-large  by  7.73 percent points, RoBERTa-base by 3.79 
  percent points, RoBERTa-large by 2.04 percents on average F1 score over 5 datasets.  
  Especially, ALBERT.xxlarge based is much more  unstable  in our senario. Thus, 
  we did multiple run with different settings, e.g.,  learning rate and batch size 
  for each dataset, and  report the  best  result  here. Comparatively,  
  RoBERTa  based is  the  most stable and has the best performance among
  all baselines and all experiment settings.

@@

  \subsubsection*{Does joint SFC improve the classfication in stage 1?}
  We  compare  the  top  1  predition and top 5 candidate labels in stage 1 from
  two-stage SFC and joint SFC in Table \ref{tbe:top1_5_accuracy}. We should note
  that  two-stage  SFC  does not optimize the stage 1 result, which is a RoBERTa
  based classification model.

  \begin{table}
    \begin{centering}
      \begin{tabular}{|c|c|c|c|c|}
        \hline
        F1 score & Two-stage SFC & Joint SFC & Gap & \tabularnewline
        \hline
        top-1 accuracy  & \textbf{81.50} & 81.07 & -0.47 & \tabularnewline
        top-5 accuracy & 94.01 & \textbf{94.30} & +0.29 & \tabularnewline
        \hline
      \end{tabular}
      \par
    \end{centering}
    \caption{The average classification accuracy in stage 1 on all five dataset.}
    \label{tbe:top1_5_accuracy}
  \end{table}

  The joint training does not improve the classification performance in stage 1,
  measured  by  top  1  accuracy.  However,  it  improves  the  quality of top 5
  candidate  class  labels.  It  can  be  explained that, a classification model
  inherently optimizes the objective loss, 0-1 error of top 1 here; the sentence
  pair similairity model in the joint SFC poses positive effect in optimizing the 
  candidate labels.

@@

  Our main
  focus  is  to  study  chatbot  building  in common situation during real-world
  application  where  only  a few sample sentences are available for each class.
  The  experimental  results  for  the  3  diverse  intention detection datasets
  (ClINC150,  BANKING77,  HWU64)  under  few-shot  settings, 5 to 20 samples per
  class,  indicate that our proposed joint SFC can achieve 2.47 percents average
  improvement   over  one  of  the  most  powerful  baseline  models,  which  is
  RoBERTa-large  classification  model.  Moreover,  the  improvement  in F1 score
  becomes  more  and  more prominent with the decrease of data sample number per
  class.

@@

  We  conduct  experiments on 5 different settings of hyperparameters, which are
  the  number of candidate class number $K$ and the number of sentence pairs $P$
  for  each  class,  on  ITG  and Amazon-670K datasets to find the most suitable
  setting  of  these  2 hyperparameters. We control the total number of sentence
  pairs  sampled for each user query into about the same level of amount(50-60).
  Through  analyzing the experimenal results shown in Table \ref{tbe:table3}, we
  find  that the best setting of hyperparameters might be different for datasets
  of  different  data  size.  The  class label number of ITG and Amazon-670k are
  similar  in  our  experimental setting, which is around 250. However the total
  amount of data points of ITG is approximately 2 times larger than Amazon-670k.
  Therefore,  the  setting of $K=5$ and $P=10$ is good enough for ITG since most
  of  the  classes  contains  over 10 data points. However, for Amazon-670k, the
  amount  of  data  sample  for  each  class  is  scarce, thus, we may need more
  candidate  classes to sample enough effective sentence pairs for sentence-pair
  model  layer  of  SFC.  In this way, the setting of $K=10$ and $P=5$ becomes a
  good choice. However, in spite of the effect on evaluation performance brought
  by different settings of hyperparameters, joint SFC still steadily outperforms
  two-stage  SFC  by  0.41  percents  in  overall  average  F1 score  on  ITG and
  Amazon-670k datasets.
